---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Alex Bastardi"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: vignette
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)
school_diversity <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-24/school_diversity.csv")

```

<br><br>

## Introduction

```{r read-examine-data}
head(school_diversity)
glimpse(school_diversity)
```

<br><br>

## Question 1

#### Is there a difference in the level of integration of schools between the school years 1994-1995 and 2016-2017?

### Methods
```{r wrangle and plot 1}
school_diversity %>%
  select(ST, SCHOOL_YEAR, diverse, int_group) %>%
  drop_na() %>%
  ggplot() +
  geom_bar() +
  aes( x= int_group, fill=SCHOOL_YEAR) +
  scale_color_brewer(palette = "Set2") + #changing color palette
  theme(legend.position = "bottom") +
  labs(x= "Level of Integration", y= "Count")-> wrangle1 #moving the legend location
print(wrangle1)

```

### Answer

Yes it appears the school year 2016-2017 has more highly integrated districts than in 1994-1995. Not integrated districts remained relatively the same between the 2 school years.

<br><br>

## Question 2

#### For school districts in NJ and NY, does the white proportion of student population differ between school years 1994-1995 and 2016-2017?

### Methods
```{r wrangle and plot 2}
school_diversity %>%
  filter( ST %in% c("NJ", "NY")) %>% #filtering states to NJ and NY
  select(ST, SCHOOL_YEAR, White) %>% #looking at only these 3 columns
  drop_na() %>%
  ggplot() +
  geom_violin(fill="grey") + #creating a violin plot
  aes(y= White, x=SCHOOL_YEAR) +
  theme_minimal() +
  labs( x=" School Year", y= "White proportion of student population")-> wrangle2
print(wrangle2)
```

### Answer

Yes, some school districts in NJ and NY had a lower white proportion of student population in 2016-2017 compared to 1994-1995.
<br><br>

## Question 3

#### Of the 6 New England states, which has the district with the highest Black porportion of student population?

### Methods
```{r wrangle and plot 3}
school_diversity %>%
  filter( ST %in% c("CT", "ME", "MA", "NH", "RI")) %>% #filtering states to New England area
  ggplot() +
  geom_boxplot(fill= "hotpink") +
  aes(x= ST, y=Black) +
  theme_minimal() +
labs(x="State", y= "Black proportion of student population")-> wrangle3 #labeling plot
print(wrangle3)
```

### Answer

A school district in Connecticut has the highest Black proportion of student population.

<br><br>

## Question 4

#### In the Philadelphia metropolitan area (NJ, DE, PA), which state has the district with the highest Asian porportion of student population?

### Methods
```{r wrangle and plot 4}
school_diversity %>%
  filter( ST %in% c("PA", "NJ", "DE")) %>% #filtering states to philly metropolitan area
  ggplot() +
  geom_boxplot(fill="blue", color="grey") + #changing color and fill to be aesthetically pleasing
  aes(x= ST, y=Asian) +
  theme_classic() +
  labs(x="State", y= "Asian proportion of student population") -> wrangle4
print(wrangle4)
```

### Answer

In the Philadelphia metropolitan area, NJ has the school district with the highest Asian proportion of student population.